---
title: "Ebola_RandomWalk"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)

library(tidyverse)
library(janitor)
library(rjags)
library(lubridate)
library(cowplot)
```
```{r}
colnames <- names(read_csv("DRC_ebola_cases.csv"))
eb <- read_csv("DRC_ebola_cases.csv", col_names = colnames, skip = 2) %>%
  mutate(report_date = mdy(report_date))
```
```{r}
eb %>% 
  ggplot(aes(x = report_date, y = confirmed_cases_change)) +
  geom_point() +
  geom_line()
```


```{r}
eb_RW <- "
model{

#Data Model
for(t in 1:n){
    y[t] ~ dnorm(x[t], tau_obs)
}

#Process Model
for(t in 2:n){
  x[t] ~ dnorm(x[t-1], tau_add)
}

#Priors
x[1] ~ dnorm(x_ic, tau_ic)
tau_obs ~ dgamma(a_obs, r_obs)
tau_add ~ dgamma(a_add, r_add)
}
"
```

```{r}
y <- eb$confirmed_cases_change + 1

data <- list(y=log(y),
             n=length(y),
             x_ic=log(1000),
             tau_ic=100,
             a_obs=1,
             r_obs=1,
             a_add=1,
             r_add=1)
```
```{r}
nchain = 3
init <- list()
for(i in 1:nchain){
  y.samp = sample(y,length(y),replace=TRUE)
  init[[i]] <- list(tau_add = 1/var(diff(log(y.samp))),
                    tau_obs = 5/var(log(y.samp)))
}
```

```{r}
j.model   <- jags.model (file = textConnection(eb_RW),
                             data = data,
                             inits = init,
                             n.chains = 3)
```


```{r, fig.asp = 1.0}
## burn-in
jags.out   <- coda.samples (model = j.model,
                            variable.names = c("tau_add","tau_obs"),
                                n.iter = 20000)
plot(jags.out)
```


```{r}
jags.out   <- coda.samples (model = j.model,
                            variable.names = c("x","tau_add","tau_obs"),
                                n.iter = 10000)
```

```{r}
#gelman.plot(jags.out)
burnin = 6000
jags.burn <- window(jags.out, start = burnin)
```

```{r}
tdate <- eb %>% 
  distinct(report_date) %>% 
  mutate(nrow = row_number()) %>% 
  mutate(nrow = as.character(nrow))

out <- as.matrix(jags.burn) %>% 
  as_tibble() %>% 
  clean_names() %>% 
  dplyr::select(-c(tau_add, tau_obs)) %>% 
  gather(x_1:x_352, key = "nrow", value = "pred") %>% 
  separate(nrow, c("delete", "nrow")) %>% 
  sample_n(1000) %>%
  left_join(tdate) %>% 
  group_by(report_date, nrow) %>% 
  summarise(cases_pred = median(pred),
            lwr_quantile = quantile(pred, 0.025),
            upr_quantile = quantile(pred, 0.975))
  

eb %>% 
  ggplot() +
  geom_point(aes(x = report_date, y = confirmed_cases_change)) +
  #geom_line(aes(x = report_date, y = confirmed_cases_change)) +
  geom_line(data = out, aes(x = report_date, y = exp(cases_pred)-1), color = "red") +
  geom_ribbon(data = out, aes(x = report_date, ymin = exp(lwr_quantile)-1, ymax = exp(upr_quantile)-1), fill = "red", alpha = 0.6) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  ylim(0, 27)
```

```{r}
data$Tmin=min(tdate$report_date)
ef.out <- ecoforecastR::fit_dlm(model=list(obs="y",fixed="~ 1 + X + Tmin"),data)
names(ef.out)
```{r, fig.asp = 1.0}
## parameter diagnostics
params <- window(ef.out$params,start=1000) ## remove burn-in
plot(params)
summary(params)
cor(as.matrix(params))
pairs(as.matrix(params))

## confidence interval
out <- as.matrix(ef.out$predict)
ci <- apply(exp(out)-1,2,quantile,c(0.025,0.5,0.975))
{plot(tdate$report_date,ci[2,])
## adjust x-axis label to be monthly if zoomed
#if(diff(time.rng) < 100){ 
#  axis.Date(1, at=seq(time[time.rng[1]],time[time.rng[2]],by='month'), #format = "%Y-%m")
#}
ecoforecastR::ciEnvelope(tdate$report_date,ci[1,],ci[3,],col=ecoforecastR::col.alpha("lightBlue",0.75))
points(tdate$report_date,y,pch="+",cex=0.5)}
``````

